\documentclass{article}

\usepackage{parskip}
\usepackage{mathtools}

\begin{document}

\title{MAT121 Exercises\\Chapter 1.2}
\author{David Huynh}
\date{}

\maketitle

\section{In Exercise 1, determine which matrices are in reduced echelon form and which others are only in echelon form}

$a=
\begin{matrix}
1 & 0 & 0 & 0\\
0 & 1 & 0 & 0\\
0 & 0 & 1 & 1
\end{matrix}
$

$b=
\begin{matrix}
1 & 0 & 1 & 0\\
0 & 1 & 1 & 0\\
0 & 0 & 0 & 1
\end{matrix}
$

$c=
\begin{matrix}
1 & 0 & 0 & 0\\
0 & 1 & 1 & 0\\
0 & 0 & 0 & 0\\
0 & 0 & 0 & 1
\end{matrix}
$

$d=
\begin{matrix}
1 & 1 & 0 & 1 & 1\\
0 & 2 & 0 & 2 & 2\\
0 & 0 & 0 & 3 & 3\\
0 & 0 & 0 & 0 & 4
\end{matrix}
$

\subsection{Solution}

Matrices a,b are in reduced echelon form.\\
Matrices d are in echelon form
Matrix c is not in echelon form, row 3 is 0-row but has non-zero row underneath

\section{Find the general solutions of the systems whose augmented matrices are given in Exercise 7}

$7)
\begin{matrix}
1 & 3 & 4 & 7\\
3 & 9 & 7 & 6
\end{matrix}
$

\subsection{Solution}

Add $-3*row1$ to row2\\
$
\begin{matrix}
1 & 3 & 4 & 7\\
0 & 0 & -5 & 15
\end{matrix}
$

Divide $row2$ by -5\\
$
\begin{matrix}
1 & 3 & 4 & 7\\
0 & 0 & 1 & 3
\end{matrix}
$

Add $-4*row2$ to row1\\
$
\begin{matrix}
1 & 3 & 0 & -5\\
0 & 0 & 1 & 3
\end{matrix}
$

$x_{1} = -5 - 3x_{2}$\\
$x_{2} = free$ because column 1 and 3 are pivot columns\\
$x_{3} = 3$

\section{In exercise 19, choose $h$ and $k$ such that the system has a: no solutions, b: a unique solution and c: many solutions. Give separate answers for each part}

\begin{align*}
x_{1} + hx_{2} &= 2\\
4x_{1} + 8x_{2} &= k 
\end{align*}

Echelon form\\
\begin{align*}
x_{1} + h & = 2\\
0x_{1} + (8-4h) & = k-8
\end{align*}

\subsection{No solution}

We must choose $h$ and $k$ such that the system's echelon form becomes inconsistent.\\
From the echelon form we can see that if we choose $k = 2$ and $h \neq 8$ we will get row with 0-variables and non-zero $b$

\subsection{Unique solution}

We must choose $h$ and $k$ such that the system's echelon form has no free variabls.\\
This happens when $h \neq 2$.

\subsection{Multiple solutions}

We must choose $h$ and $k$ such that the system's echelol form has free variables and is consistent.\\
This happens when $h = 2$ and $k = 8$

\section{In exercise 21 mark each statement True or False. Justify each answer.}

a)\textbf{In some cases, a matrix may be row reduced to more than one matrix in reduced echelon form, using different sequences of row operations.}\\
False

b)\textbf{The row reduction algorithm applies only to augmented matrices for a linear system.}\\
False, it can be applied on non-augmented matrices too

c) \textbf{A basic variable in a linear system is a variable that corresponds to a pivot column in the coefficient matrix.}\\
True, per definition

d) \textbf{Finding a parametric description of the solution set of a linear system is the same as \emph{solving} the system.}\\
True, parametric description expresses the variables as combinations of eachother or free.

e) \textbf{If one row in an echelon form of an augmented matrix is $[0 0 0 5 0]$, then the associated linear system is inconsistent.}\\
False, in this case the rightmost column has to be non-zero, the others zero
 
\end{document}
